---
title: "HW 1, Recreating parts of the 538 Bechdel Analysis"
author: "Andrew Kotarski"
output: github_document
---

# Introduction

In this mini analysis we work with the data used in the FiveThirtyEight story titled ["The Dollar-And-Cents Case Against Hollywood's Exclusion of Women"](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/).
Your task is to fill in the blanks denoted by `___`.

The point of this first homework assignment is not only to complete an analysis, but also
to have you explore running `R` code within the RMarkdown framework. 

I have the .Rmd file set up to generate a markdown file within the project directory when you `knit`, and you should be able to `knit` as soon as you install the `FiveThirtyEight` package. Run `install.packages("fivethirtyeight")` in the console before knitting for the first time.  

## Some tips and reminders:

`R` code is run either within `R` code "chunks"

```{r}
# R code goes here. You can run this code chunk by pressing the green arrow in the 
# top-right of the code chunk, or go to Code --> Run... to see various keyboard shortcuts
print("R code within this chunk is evaluated top-to-bottom!")
```
   
or "in-line" when you knit the document with markdown text like this: 
`r "Wow I can also run R code in-line!"`
And like this: 2+2 = `r 2+2`.  

When you `knit` the RMarkdown document (press the `knit` button at the top of the screen, or 
click `File` $\rightarrow$ `Knit Document`) `R` will compile your text, while also running
all of the code, from top-to-bottom. 

Also note that I can add "comments" to RMarkdown text
<!-- like this -->
and they do not get evaluated. 



# Analyses

## Data and packages

We start with loading the packages we'll use.

```{r load-packages, message=FALSE}
# remember to run install.packages("fivethirtyeight") in the console 
# before calling library. You only have to install once, but you need to use library()
# every time at the start of an RMarkdown document calling the functions/data from the 
# package
library(fivethirtyeight)
library(tidyverse)
```

The dataset contains information on `r nrow(bechdel)` movies released between `r min(bechdel$year)` and `r max(bechdel$year)`.
However we'll focus our analysis on movies released between 1990 and 2013.

```{r }
bechdel90_13 <- bechdel %>%  
  filter(between(year, 1990, 2013)) 
# filters the column year between 1990 and 2013
nrow(bechdel90_13)
```

<!-- (try and answer this one using "in-line" R code). See line 56 above  -->
 
There are `r nrow(bechdel90_13)` such movies. 

The financial variables we'll focus on are the following:

-   `budget_2013`: Budget in 2013 inflation adjusted dollars
-   `domgross_2013`: Domestic gross (US) in 2013 inflation adjusted dollars
-   `intgross_2013`: Total International (i.e., worldwide) gross in 2013 inflation adjusted dollars

And we'll also use the `binary` and `clean_test` variables for **grouping**.

## Analysis

Let's take a look at how median budget and gross vary by whether the movie passed the Bechdel test, which is stored in the `binary` variable.

```{r message = FALSE}
bechdel90_13 %>%
  group_by(binary) %>% # groups data by the unique `binary` variables 
  # the code below creates summaries of the various columns called 
  # according to the function used. In this case, `median()`
  summarize(med_budget = median(budget_2013), 
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
# Note that many of the functions in the tidyverse follow both American and British
# spelling conventions, so you may see this function as "summarise" online or in 
# some class material 
```

Next, let's take a look at how median budget and gross vary by a more detailed indicator of the Bechdel test result.
This information is stored in the `clean_test` variable, which takes on the following values:

-   `ok` = passes test
-   `dubious`
-   `men` = women only talk about men
-   `notalk` = women don't talk to each other
-   `nowomen` = fewer than two women


<!-- Fill in the group_by line below and remove the `#` so it is no longer commented out.  -->
<!-- Hint: What were we interested in comparing across in the R chunk above, and  -->
<!-- what did we group_by then? -->

```{r message = FALSE}
bechdel90_13 %>%
  group_by(`clean_test`) %>%
  summarize(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

In order to evaluate how return on investment varies among movies that pass and fail the Bechdel test, we'll first create a new variable called `roi` as the ratio of the gross to budget.

```{r}
bechdel90_13 <- bechdel90_13 %>%
  mutate(roi = (intgross_2013 + domgross_2013) / budget_2013)
```

Let's see which movies have the highest return on investment.

```{r}
bechdel90_13 %>%
  arrange(desc(roi)) %>% # arranges data in descending order according to `roi` column
  select(title, roi, year) # selects the columns title, roi, year
```

Below is a visualization of the return on investment by test result, however it's difficult to see the distributions due to a few extreme observations.

```{r warning = FALSE}
# we will cover the various ggplot calls in class soon! 
ggplot(data = bechdel90_13, 
       mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       x = "Detailed Bechdel result",
       y = "Return on investment",
       color = "Binary Bechdel result")
```

What are those movies with *very* high returns on investment?

```{r}
bechdel90_13 %>%
  filter(roi > 400) %>% 
  select(title, budget_2013, domgross_2013, year) 

```

Zooming in on the movies with `roi < 16` provides a better view of how the medians across the categories compare:

<!-- Hint to answer the above: look at the code chunk below after running it...  -->
<!-- which line do you think controls the limits of the Y axis? -->

```{r warning = FALSE}
ggplot(data = bechdel90_13, mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       subtitle = "View on medians", # Something about zooming in to a certain level
       x = "Detailed Bechdel result",
       y = "Return on investment",
       color = "Binary Bechdel result") +
  coord_cartesian(ylim = c(0, 15)) 
```


When you have finished filling in all of the `___`'s make sure to knit the document a final time, commit your changes with git (we will review this in class and practice more on Wednesday Jan 26) and push your changes to Github (see Thursday's lab for more information). 



