---
title: "Final Project"
author: "Group 6"
date: "4/2/2022"
output: html_document
---

## Final project Setup

```{r load-tidyverse-and-dataset, include=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
airplane <- read_csv("data/Airplane_Crashes_and_Fatalities_Since_1908.csv")
```

Change variable tyoes:

```{r change-variable-types}
airplane$date <- lubridate::mdy(airplane$Date)
airplane$aboard <- as.numeric(airplane$Aboard)
```


## Exploring the dataset

Crashes per day as scatterplot:
```{r crashes-per-day-plot}
airplane %>%
  count(date) %>%
  ggplot(aes(x=date,y=n)) +
  geom_point()
```

Crashes per day in table in descending order of crashes:
```{r crashes-per-day-tibble}
airplane %>%
  count(date) %>%
  arrange(desc(n))
```

Specific observations on 08/31/1988:
```{r 1988-08-31, echo=FALSE}
airplane %>%
  filter(date=="1988-08-31")
```
Specific observations on 09/11/2001:
```{r 2001-09-11, echo=FALSE}
airplane %>%
  filter(date=="2001-09-11")
```

Crashes by operator in descending order:
```{r operators, echo=FALSE}
airplane %>%
  count(airplane$Operator) %>%
  arrange(desc(n))
```

Crashes by aircraft type in descending order:
```{r 1944-06-06, echo=FALSE}
airplane %>%
  count(airplane$`AC Type`) %>%
  arrange(desc(n))
```

Scatterplot of crashes per year:
```{r years-grouped}
airplane %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  count() %>%
  ggplot(aes(x=year,y=n)) +
  geom_point()
```

Scatterplot of mean fatality proportion by year: 
```{r fatality-proportion-plot}
airplane<- airplane %>%
  mutate(fatality_prop = Fatalities / aboard)
airplane %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(mean=mean(fatality_prop)) %>%
  ggplot(aes(x=year,y=mean)) +
  geom_point() +
  geom_smooth()
```

```{r Greatest fatalities by Operator, Location, Time, Date }
airplane %>%
  group_by(Operator, Location, Time, Date ) %>%
summarise(Fatalities) %>%
 arrange(desc(Fatalities)) 
```
```{r visulation of fatalities}
ggplot(data = airplane, mapping = aes(x = Fatalities)) +
  geom_histogram(bins = 50)
```

```{r summary stats for fatalities}
airplane %>%
  summarise(
            mean = mean(Fatalities, na.rm = TRUE),
            med = median(Fatalities, na.rm = TRUE),
            sd = sd(Fatalities, na.rm = TRUE), 
            q1 = quantile(Fatalities, na.rm = TRUE),
            q3 = quantile(Fatalities,na.rm = TRUE))
```


```{r fatalities-years-grouped}
airplane %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(sum=sum(Fatalities)) %>%
  ggplot(aes(x=year,y=sum)) +
  geom_point() +
  geom_smooth(method="loess") +
  labs(x="Year",y="Fatalities per year")
```


```{r crashes-vs-fatalities}
crashes_by_year <- airplane %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  count() 
fatalities_by_year <- airplane %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(fatalities=sum(Fatalities))
crashes_and_fatalities <- crashes_by_year %>%
  left_join(fatalities_by_year, by="year")
ggplot(crashes_and_fatalities,aes(x=n,y=fatalities)) +
  geom_point()
```