---
title: "Lab 03 - Nobel laureates"
author: "Team 6"
date: "2/16/22"
output: github_document
---

### Load packages and data

```{r load_tidyverse, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are `r nrow(nobel)` observations and `r ncol(nobel)` variables in the dataset. Each row represents a laureate.

### Exercise 2

```{r new-data-frame}
nobel_living <- nobel %>%
  filter(!is.na(country)) %>%
  filter(gender!="org") %>%
  filter(is.na(died_date))
```
The data frame nobel_living contains `r nrow(nobel_living)` observations. 

### Modify nobel_living dataframe

```{r US-nobel-laureates}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r nobel-living-science}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```
The data frame nobel_living_science contains Nobel laureates in the areas of
physics, medicine, chemistry and economics classified into "US" or "other".

### Exercise 3

Parts of the Buzzfeed headline are supported by the data. There were more Nobel
laureates based inside the USA than outside the USA when they won their 
prizes, however, we cannot tell from this visualization whether or not the laureates immigrated to the USA or were born here.

```{r barplot}
ggplot(nobel_living_science, 
       aes(x= country_us)) +
  geom_bar()+
  coord_flip()+
  facet_wrap(~category) +
  labs(y="Count",x="Country of laureate in prize year")
```

### Exercise 4

```{r create-born-country-variable}
nobel_living_science <- nobel_living_science %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other"))
```
`r sum(nobel_living_science$born_country_us=="USA")` of the living laureates in the categories Physics, Medicine, Chemistry and Economics were born in the US.

### Exercise 5

```{r barplot}
ggplot(nobel_living_science, 
       aes(x= country_us,
           fill = born_country_us)) +
  geom_bar(position = "fill")+
  coord_flip()+
  facet_wrap(~category) +
  
  labs(y="Count",x="Country of laureate in prize year", fill = "Born in US")
```
I would agree with buzzfeeds claim that "of those US-based Nobel laureates, many were born in other countries." While there isint a clear definition on the numeric value of "many" there seems to be supporting evidence that of the US-based Nobel Laureates a fair portion were born in other countries. 

### Exercise 6

...
